\chapter*{Abstract}

Transportation remains one of the largest contributors to global greenhouse gas emissions. Data-driven sustainability analyses are therefore required to quantify transportation emissions and evaluate the environmental performance of transportation processes. While Object-Centric Process Mining (OCPM) provides suitable techniques for analyzing complex transportation processes involving multiple interacting entities, the availability of publicly available Object-Centric Event Logs (OCELs) containing sustainability-relevant transportation data remains limited.

This work investigates the suitability of General Transit Feed Specification (GTFS) data for sustainability analysis in OCPM. We show that GTFS provides the information required for transportation emission estimation, including travel durations, traveled distances, vehicle movements, delays, GPS coordinates, and occupancy information. Based on these data, we present GTFS2OCEL, a construction pipeline that transforms GTFS Static and GTFS-Realtime data into the OCEL 2.0 standard by deriving events, objects, relationships, and sustainability-relevant attributes. Furthermore, we provide the generated sustainability-aware OCELs as publicly available datasets for future process mining and sustainability research.

To evaluate the generated event logs, we investigate whether travel duration can be used as a substitute for traveled distance in transportation emission estimation and quantify the resulting approximation error at different levels of granularity. The results show that duration-based emission estimation is feasible but introduces approximation errors, which become more pronounced at finer levels of analysis. Despite limitations caused by missing information, such as detailed vehicle characteristics, GTFS2OCEL demonstrates that GTFS data can be systematically transformed into sustainability-aware object-centric event logs, providing a valuable foundation for transportation sustainability analysis and future OCPM research.